Data description

### 2017-02-09 Data/data_explnation.Rmd
update explnation.Rmd

__Interactions:__ This include `transect`, `visitation` and `out` data. Data can be used pooled (2067 interactions (5732 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`. 
```{r}
int <- read.csv("interactions.csv", h = T)
head(int)
nrow(int)
sum(int$Frequency, na.rm = TRUE)
```
Challenges:  
- long tail of singletones  
- not all species are taxonomically resolved. What to do with specimens we ony know Genus?  
- Low sample size when you split per site and round

@@

>>>>

__Interactions:__ This include `transect`, `visitation` and `out` data. Data can be used pooled (2067 interactions (5732 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`.

```{r}
int <- read.csv("interactions.csv", h = T)
head(int)
nrow(int)
sum(int$Frequency, na.rm = TRUE)
```
Challenges:  
- long tail of singletones  
- not all species are taxonomically resolved. What to do with specimens we ony know Genus?  
- Low sample size when you split per site and round

@@

# Reshaping the dataset as input of the `as.alienData()` function

**Extract from the roxygen2 documentation:**

```
#' @param idObs A data.frame which is mandatory and will help to check consistency and prevent errors among unique identifiers of each alienData arguments. The first column (idSite) contains unique identifier of where the observation was made. The second column (idTime) is not mandatory and contains temporal information: an unique identifier at the time the sample has been taken (needed for timeseries analysis). The third column (idSpcies) is an unique identifier of the species sampled at time (idTime) and location (idSite). The fourth column is an unique identifier of individu of species (idSp) observed at time (idTime) and location (idSite).
#' @param interactPair A data.frame which contains interaction at the finest level (individus or species). The first two columns are idFrom and idTo and determine the sens of the interaction. idFrom and idTo are unique identifier of species or individu documented in the idObs data.frame. Finaly, the thrid column is the strength of the interaction (Please see details).
#' @param coOcc A square symmetric matrix of 0s and 1s that define co-occurence patterns among pairs of species. If this matrix is not provided, the co-occurence matrix is derived from the coAbund matrix else the interactSp matrix (see return section).
#' @param coAbund A square symmetric matrix that includes any types of values, defining co-abundance patterns among pairs of species. TODO: Not implemented yet.
#' @param siteEnv A matrix or a data.frame where each column is a descriptor of the sites. TODO: siteEnv should cover the possibility that environmental variables could be taken at several times.
#' @param traitSp A matrix or a data.frame where each column is a trait characterizing all species. The first column is a unique identifier of the species documented in idObs data.frame.
#' @param traitInd A matrix or a data.frame where each column is a trait characterizing an individual. The first column is a unique identifier of the individu documented in idObs data.frame.
#' @param phylo A square symmetric matrix describing the phylogenetic relationships between pairs of all species (see details). TODO: Not implemented yet.
#' @param scaleSiteEnv Logical. Whether the columns of X should be centred and divided by the standard deviation. Default is TRUE.
#' @param scaleTrait Logical. Whether the rows of Tr should be centred and divided by the standard deviation. Default is TRUE.
#' @param interceptSiteEnv Logical. Whether a column of 1s should be added to X. Default is TRUE.
#' @param interceptTrait Logical. Whether a row of 1s should be added to Tr. Default is TRUE.
#'
#' @details
#'
#' The strength of the interactions defined in the third column of \code{interactPair} can be a 0 if no direct interaction has been observed (defined as true absence of interaction) or any numerical value. Undocumented interactions among species or individus will be assumed as NA by default.
```

```{r, echo=FALSE}
# Create dir for formatted data
dir.create('./Data/alienData', showWarnings = FALSE)
```

### Each section below transform your data into as.alienData args

#### Get idObs

```{r}
# load data
int <- read.csv("interactions.csv", h = T, stringsAsFactors=FALSE)
flw <- read.csv("flowers.csv", h = T, stringsAsFactors=FALSE)
plant_tr <- read.csv("traits_plants.csv", h = T, stringsAsFactors=FALSE)
pol_tr <- read.csv("traits_pollinators.csv", h = T, stringsAsFactors=FALSE)

# Create unique IDs for pollinator individus
int$idPol <- seq(1,nrow(int),1)
# Create unique IDs for plant individus
int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1)

# subset pollinator and plant with their own ids
pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")])
colnames(pol) <- c("idSite","idTime","idSp","idInd")
plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")])
colnames(plant) <- c("idSite","idTime","idSp","idInd")

# Set idObs by merging both dfs
idObs <- rbind(pol,plant)

#Turn all columns as factor
idObs <- as.data.frame(lapply(idObs,as.factor))
```

#### Get InteracPair

```{r}
interactPair <- int[,c("idPol","idPlant","Frequency")]
```

#### traitInd

Trait by individus

```{r}
traitInd <- int[,c("idPol","Pollinator_sex")]
traitInd$traitName <- "sex"
traitInd <- traitInd[,c('idPol','traitName','Pollinator_sex')]
names(traitInd)[3] <- 'value'
```

#### traitSp

Trait by species
Note Removed family as a trait

```{r}
library(reshape2)

traitSpPlant <- melt(plant_tr[,-2],id=c("Plant_gen_sp"),value.name="value",variable.name="traitName")
names(traitSpPlant)[1] <- "idSp"

traitSpPol <- melt(pol_tr[,-4],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName")
names(traitSpPol)[1] <- "idSp"

traitSp <- rbind(traitSpPlant,traitSpPol)
traitSp$idSp <- as.factor(traitSp$idSp)

#### errors
# Some entries in traitSp (merge of pol_tr and plant_tr) are documenting species which are not present in `ìnt`, the interaction pairwise data.frame.
removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
traitSp <- subset(traitSp, !idSp %in% removeSpecies )
traitSp$idSp <- droplevels(traitSp$idSp)
```

#### Save all args of the as.alienData() into a RData file.

```{r}
save(idObs,interactPair,traitSp,traitInd,file="./alienData/argsAlienData.RData")
```
